Avoiding tipping points in the management of ecological systems: a non-parametric Bayesian approach

### 2013-11-20 inst/doc/nonparametric-bayes.tex
writing

\begin{table}[ht]
\begin{center}
\begin{tabular}{rrrr}
  \hline
 & Allen & Ricker & Myers \\ 
  \hline
DIC & 50.14 & 49.45 & 50.61 \\ 
  AIC & -24.60 & -30.07 & -27.19 \\ 
  BIC & -17.85 & -25.00 & -20.44 \\ 
   \hline
\end{tabular}
\end{center}
\end{table}

@@

\section{Appendix}\label{appendix}

\subsection{Model definitions and
estimation}\label{model-definitions-and-estimation}

\[X_{t+1} = Z_t X_t e^{r \left(1 - \frac{S_t}{K} \right) } \label{rickereq} \]

\begin{figure}[htbp]
\centering
\includegraphics{figure/nonparametric-bayes-unnamed-chunk-4.pdf}
\caption{Traces from the MCMC estimate of the Ricker model}
\end{figure}

\begin{figure}[htbp]
\centering
\includegraphics{figure/nonparametric-bayes-unnamed-chunk-5.pdf}
\caption{Posteriors from the MCMC estimate of the Ricker model}
\end{figure}

@@

\begin{enumerate}
\def\labelenumi{(\arabic{enumi})}
\itemsep1pt\parskip0pt\parsep0pt
\item
  \[ X_{t+1} = Z_t \frac{r S_t^{\theta}}{1 - \frac{S_t^\theta}{K}}\]
\end{enumerate}

\begin{figure}[htbp]
\centering
\includegraphics{figure/nonparametric-bayes-unnamed-chunk-7.pdf}
\caption{Traces from the MCMC estimate of the Myers model}
\end{figure}

\begin{figure}[htbp]
\centering
\includegraphics{figure/nonparametric-bayes-unnamed-chunk-8.pdf}
\caption{Posterior distributions from the MCMC estimates of the Myers
model}
\end{figure}

@@

Eq S3: Allen model

\begin{enumerate}
\def\labelenumi{(\arabic{enumi})}
\setcounter{enumi}{1}
\itemsep1pt\parskip0pt\parsep0pt
\item
  \[f(S_t) = S_t e^{r \left(1 - \frac{S_t}{K}\right)\left(S_t - C\right)} \]
\end{enumerate}

Figure S3: Allen model: prior and posterior distributions for parameter
estimates.

\begin{figure}[htbp]
\centering
\includegraphics{figure/nonparametric-bayes-unnamed-chunk-10.pdf}
\caption{Traces from the MCMC estimate of the Allen model}
\end{figure}

\begin{figure}[htbp]
\centering
\includegraphics{figure/nonparametric-bayes-unnamed-chunk-11.pdf}
\caption{Posteriors from the MCMC estimate of the Allen model}
\end{figure}

\begin{verbatim}
Error: object 'allen_priors_xtable' not found
\end{verbatim}

@@

\begin{figure}[htbp]
\centering
\includegraphics{figure/nonparametric-bayes-unnamed-chunk-13.pdf}
\caption{Traces from the MCMC estimates of the GP model}
\end{figure}

\begin{figure}[htbp]
\centering
\includegraphics{figure/nonparametric-bayes-unnamed-chunk-14.pdf}
\caption{Posterior distributions from the MCMC estimate of the GP model.
Prior curves shown in red.}
\end{figure}

@@

\subsection{Optimal Control Problem}\label{optimal-control-problem}

@@

\subsubsection{Training data}\label{training-data}

Eacho of our models $f(S_t)$ must be estimated from training data, which
we simulate from the Allen model with parameters $r = $
\texttt{r p{[}1{]}}, $K =$ \texttt{r p{[}2{]}}, $C =$
\texttt{r p{[}3{]}}, and $\sigma_g =$ \texttt{r sigma\_g} for $T=$ 40
timesteps, starting at initial condition $X_0 = $ 5.5. The training data
can be seen in Figure 1.

\subsubsection{Sensitivity Analysis}\label{sensitivity-analysis}

@@

>>>>

\begin{table}[ht]
\begin{center}
\begin{tabular}{rrrr}
  \hline
 & Allen & Ricker & Myers \\ 
  \hline
DIC & 50.14 & 49.45 & 50.61 \\ 
  AIC & -24.60 & -30.07 & -27.19 \\ 
  BIC & -17.85 & -25.00 & -20.44 \\ 
   \hline
\end{tabular}
\caption{Model choice scores for several common criteria all (wrongly) select the simplest model. As the true (Allen) model is not distinguishable from the simpler (Ricker) model in the region of the observed data, this error cannot be avoided regardless of the model choice criterion. This highlights the danger of model choice when the selected model will be used outside of the observed range of the data.}
\end{center}
\end{table}

@@

\section{Model definitions and
estimation}\label{model-definitions-and-estimation}

\subsection{Ricker Model}\label{ricker-model}

The Ricker model is given by

\begin{enumerate}
\def\labelenumi{(\arabic{enumi})}
\itemsep1pt\parskip0pt\parsep0pt
\item
  \[X_{t+1} = Z_t X_t e^{r \left(1 - \frac{S_t}{K} \right) } \]
\end{enumerate}

where $Z_t$ is log-normal noise of mean unity and log standard deviation
$\sigma$, representing the stochastic growth, $X_t$ the stock size at
time $t$, $S_t$ the escapement (unharvested population that will recruit
in the following year, $S_t = X_t - h_t$). We place uniform priors on
the growth rate $r$, carrying capacity $K$, and log-normal standard
deviation parameter $\sigma$, over ranges given in Table 1. Posteriors
are inferred by Gibbs sampling using Jags ({\textbf{???}}) (see provided
code).

\begin{figure}[htbp]
\centering
\includegraphics{figure/nonparametric-bayes-unnamed-chunk-2.pdf}
\caption{Posteriors from the MCMC estimate of the Ricker model}
\end{figure}

@@

\subsection{Myers Model}\label{myers-model}

The Myers model (Myers et al. 1995) is given by

\begin{enumerate}
\def\labelenumi{(\arabic{enumi})}
\setcounter{enumi}{1}
\itemsep1pt\parskip0pt\parsep0pt
\item
  \[ X_{t+1} = Z_t \frac{r S_t^{\theta}}{1 - \frac{S_t^\theta}{K}}\]
\end{enumerate}

where $Z_t$ is log-normal noise of mean unity and log standard deviation
$\sigma$, representing the stochastic growth, $X_t$ the stock size at
time $t$, $S_t$ the escapement (unharvested population that will recruit
in the following year, $S_t = X_t - h_t$). We place uniform priors on
the growth rate $r$, carrying capacity $K$, $\theta$ controls the
strength of the nonlinearity, exhibiting an allee effect for
$\theta \geq 2$, and log-normal standard deviation parameter $\sigma$,
over ranges given in Table 1. Posteriors are inferred by Gibbs sampling
using Jags ({\textbf{???}}) (see code provided).

\begin{figure}[htbp]
\centering
\includegraphics{figure/nonparametric-bayes-unnamed-chunk-4.pdf}
\caption{Traces from the MCMC estimate of the Myers model}
\end{figure}

\begin{figure}[htbp]
\centering
\includegraphics{figure/nonparametric-bayes-unnamed-chunk-5.pdf}
\caption{Posterior distributions from the MCMC estimates of the Myers
model}
\end{figure}

@@

\subsection{Allen model}\label{allen-model}

The Allen model (Allen and Tanner 2005) is given by

\begin{enumerate}
\def\labelenumi{(\arabic{enumi})}
\setcounter{enumi}{2}
\itemsep1pt\parskip0pt\parsep0pt
\item
  \[f(S_t) = S_t e^{r \left(1 - \frac{S_t}{K}\right)\left(S_t - C\right)} \]
\end{enumerate}

where $Z_t$ is log-normal noise of mean unity and log standard deviation
$\sigma$, representing the stochastic growth, $X_t$ the stock size at
time $t$, $S_t$ the escapement (unharvested population that will recruit
in the following year, $S_t = X_t - h_t$). We place uniform priors on
the growth rate $r$, carrying capacity $K$, allee threshold $C$, and
log-normal standard deviation parameter $\sigma$, over ranges given in
Table 1. Posteriors are inferred by Gibbs sampling using Jags
({\textbf{???}}) (see code provided).

\begin{figure}[htbp]
\centering
\includegraphics{figure/nonparametric-bayes-unnamed-chunk-7.pdf}
\caption{Traces from the MCMC estimate of the Allen model}
\end{figure}

\begin{figure}[htbp]
\centering
\includegraphics{figure/nonparametric-bayes-unnamed-chunk-8.pdf}
\caption{Posteriors from the MCMC estimate of the Allen model}
\end{figure}

\begin{table}[ht]
\begin{center}
\begin{tabular}{rlrr}
  \hline
 & parameter & lower.bound & upper.bound \\ 
  \hline
1 & r0 & 0.01 & 6.00 \\ 
  2 & K & 0.01 & 20.00 \\ 
  3 & theta & 0.01 & 20.00 \\ 
  4 & sigma & 0.00 & 100.00 \\ 
   \hline
\end{tabular}
\caption{parameterization range for the uniform priors in the Allen model}
\end{center}
\end{table}

@@

\begin{figure}[htbp]
\centering
\includegraphics{figure/nonparametric-bayes-unnamed-chunk-10.pdf}
\caption{Traces from the MCMC estimates of the GP model}
\end{figure}

\begin{figure}[htbp]
\centering
\includegraphics{figure/nonparametric-bayes-unnamed-chunk-11.pdf}
\caption{Posterior distributions from the MCMC estimate of the GP model.
Prior curves shown in red.}
\end{figure}

@@

\subsection{Optimal Control Problem
Specification}\label{optimal-control-problem-specification}

@@

\subsection{Training data}\label{training-data}

Eacho of our models $f(S_t)$ must be estimated from training data, which
we simulate from the Allen model with parameters $r = $ 2, $K =$ 8,
$C =$ 5, and $\sigma_g =$ 0.05 for $T=$ 40 timesteps, starting at
initial condition $X_0 = $ 5.5. The training data can be seen in Figure
1.

\subsection{Code}\label{code}

A copy of the script to reconstruct the simulations and analysis shown
here is provided in the supplemental materials, and through this
version-stable link to the project's Github code repostitory,
\href{}{nonparameteric-bayes.R}. This code is dynamically embedded into
the manuscript using Knitr, (Xie 2012). The script relies on custom
routines for executing the estimation of the Gaussian process and the
for solving the stochastic dynamic programming problem. These routines
are provided as an R package, \href{}{nonparameteric-bayes}, also
available on Github.

\subsection{Sensitivity Analysis}\label{sensitivity-analysis}

@@

Xie, Yihui. 2012. \emph{Knitr: A General-purpose Package for Dynamic
Report Generation in R}. \url{http://yihui.name/knitr/}.
